```{r db-connections, include = FALSE}
eval_connections <- FALSE
if(Sys.getenv("GLOBAL_EVAL") != "") eval_connections <- Sys.getenv("GLOBAL_EVAL")
```

```{r, eval = TRUE, include = FALSE}
library(DBI)
library(odbc)
library(config)
library(keyring)
```

# Introduction to database connections

## Connecting via DSN
*Connect using defined Data Source Name (DSN)*

1. Load the `DBI` and `odbc` packages
    ```{r, eval = TRUE}
    library(DBI)
    library(odbc)
    ```

1. Use `odbcListDatasources` to list available DSNs
    ```{r, eval = TRUE}
    odbcListDataSources()
    ```
    
1. Use `dbConnect` to connect to a database using the `odbc` function and a DSN
    ```{r, eval = eval_connections}
    {{dsn_connection}}
    ```
    
1. Disconnect using `dbDisconnect`
    ```{r, eval = eval_connections}
    dbDisconnect(con)
    ```

## Connect with a connection string
*Connect by specifying all connection details in `dbConnect`*

1. Use `dbConnect` and `odbc` to connect to a database, but this time all connection details are provided
    ```{r, eval = eval_connections}
    {{connection_string_connection}}
    ```
    
1. Disconnect using `dbDisconnect`
    ```{r, eval = eval_connections}
    dbDisconnect(con)
    ```
    
## Secure connection details
*Use `config` and `keyring` to secure connection credentials*

1. Load the `config` package
    ```{r, eval = eval_connections}
    library(config)
    ```
    
1. Get the current config using the `get` function and store the results in an object called `config`
    ```{r, eval = eval_connections}
    config <- get()
    ```

1. Use `str` to investigate the contents of `config`
    ```{r, eval = eval_connections}
    str(config)
    ```

1. Connect using details provided in `config`
    ```{r, eval = eval_connections}
    con <- dbConnect(odbc(),
                     driver = config$driver,
                     host = config$host,
                     user = config$user,
                     pwd = config$pwd,
                     port = config$port,
                     database = config$dbname,
                     bigint = config$bigint)
    ```
    
1. Disconnect using `dbDisconnect`
    ```{r, eval = eval_connections}
    dbDisconnect(con)
    ```
    
1. Load the `keyring` package
    ```{r, eval = eval_connections}
    library(keyring)
    ```

1. Store the database host, username, and password using `keyring`
    ```{r, eval = eval_connections}
    key_set("postgres", "host")
    key_set("postgres", "username")
    key_set("postgres", "password")
    ```
    
1. Use the stored credentials along with `dbConnect` to connect to the database
    ```{r, eval = eval_connections}
    con <- dbConnect(odbc(),
                     driver = "postgresql",
                     host = key_get("postgres", "host"),
                     user = key_get("postgres", "username"),
                     pwd = key_get("postgres", "password"),
                     port = 5432,
                     database = "postgres",
                     bigint = "integer")
    ```

1. Discnonect using `dbDisconnect`
    ```{r, eval = eval_connections}
    dbDisconnect(con)
    ```

